---
title: "Data Tidying"
author: "Lia Domke"
date: "11/6/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
# loading in packages
library(dplyr)
library(tidyr)
```

# Outline

# Read and clean data
```{r}
catch_original <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", 
                           stringsAsFactors = FALSE)

# if you're on a windows you may need to use the below syntax

# catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", 
#                               method = "libcurl"),
#                    stringsAsFactors = FALSE)

head(catch_original)
# keyboard shortcut for pipe = command + shift + m %>% 
```

Remove `all` column and `notes` column because they do not contribute any additional information to the dataframe. 
```{r}
catch_data <- catch_original %>%  # the pipe operater always puts the original df into the first argument, can specify others
  dplyr::select(-All, -notesRegCode) # can also use select to tell R what columns to keep (w/o -) 

head(catch_data)
```
Quality check of the data
```{r}
summary(catch_data) # note R classifies Chinook as a character NOT numeric. 
```
## Fix Chinook column

Use `mutate` to fix Chinook column

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook)) # name of column you're creating or overwriting is FIRST

# "NAs introduced by coercion" means that R got to a character it didn't know how to turn into a numeric, made it a NA

summary(catch_clean)
```
Finding the rows that got turned into NA
```{r}
i <- which(is.na(catch_clean$Chinook))
i
```
Look at row with Chinook error
```{r}
catch_original[i, ] # value of Chinook is an uppercase i 
```

# Reshape data

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), # can use tidyverse info so can say pivot all BUT
               names_to = "species", 
               values_to = "count") 

head(catch_long) # 4 columns instead of 7 but is longer than it was before
```

```{r}
catch_wide <- catch_long %>% 
  pivot_wider(names_from = Year, 
              values_from = count)
head(catch_wide)
```

`rename` count column to `catch_thousands`

```{r}
catch_long <- catch_long %>% 
  rename(catch_thousands = count) %>% # new column 
  mutate(catch = catch_thousands * 1000) %>%
  dplyr::select(-catch_thousands)
  

head(catch_long)
```

# Summarize data
```{r}
mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarize(catch_mean = mean(catch), 
            num_obs = n()) # summarize will do the apply and combine

mean_region

catch_by_species <- catch_long %>% 
  group_by(species) %>%
  summarise(catch_sum = sum(catch)) %>% 
  arrange(desc(catch_sum))

catch_by_species
```

Total species pre 1900

```{r}
total_species_pre_1900 <- catch_long %>% 
  filter(Year < 1900) %>% 
  group_by(species) %>%
  summarise(catch_sum = sum(catch)) %>% 
  arrange(desc(catch_sum))

total_species_pre_1900
```

# Join to Region Table

Read in region table
```{r}
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", 
                        stringsAsFactors = F) %>% 
  select(code, mgmtArea)

head(region_defs)
```

```{r}
# catch_joined <- left_join(catch_long, region_defs, by = "Region") # this doesn't work because the names aren't the same in each dataframe

# trimws - removes the white space at the end of words to make joins work better 
# tolower - makes character vectors to lowercase
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))
```

# Seperate and Unite

```{r}
site_codes <- data.frame(site = c("HAW-100",
                                  "HAW-101", 
                                  "OAH-102",
                                  "OAH-103",
                                  "MAI-100"),
                         stringsAsFactors = F) # these sites codes have multiple pieces of information cant group over island or site nuumber because its stuck together. 
```

Separate island and site number using `separate`

```{r}
site_codes_split <- site_codes %>% 
  separate(site, into = c("island", "site_num"), sep = "-", 
           remove = F) # needs to have consistent seperator. In this case dash
site_codes_split
```

```{r}
united <- site_codes_split %>% 
  unite(site_code, island, site_num, sep = "_")

united
```

