---
title: "functions"
author: "Lia Domke"
date: "11/6/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(roxygen2)
```

Functions

If you're copying and pasting R code within or between scripts you should consider writing generalized functions that can be used over and over again throughout the code or your general work. Remember: "Do not repeat yourself" or DRY concept

```{r}
airtemps = c(212, 30, 78, 32) #including some known datapoints to make sure the function works appropriately
# if you wanted celsius for all of these you might have to run this multiple times
celsius1 <- (airtemps[1] - 32) * 5/9
celsius2 <- (airtemps[2] - 32) * 5/9
celsius3 <- (airtemps[3] - 32) * 5/9
celsius4 <- (airtemps[4] - 32) * 5/9
```

Create a function to converte farenheit to celsius

Define a set of code you want to reuse and abstracting away what you are repeating 
fucntions will work on scalers and vectors and will do the iterations (that for loops will do too). 
For-loops can be useful because you can see the intermediaries and can help de-bug. 
```{r}
fahr2celsius <- function(fahr) {
  celsius <- (fahr - 32) * 5/9 # this is the formula for conversions
  return(celsius) # tells R to split this result out
} # does this silently in the background but the function does exist in the environment
```

Use the function
```{r}
celsius_f <- fahr2celsius(airtemps[1])
celsius_f == celsius1 # do these match?

celsius <- fahr2celsius(airtemps)
celsius4 == celsius[4]
```

Function to calculate farenheit from celsius
```{r}
#' converts celsius to fahrenheit temperatures
#'
#' a little needed but more detailed explanation
#' goes here.
#' @param celsius the temperature in degrees celsius to be converted
#' @return the temperature in fahrenheit

celsius2fahr <- function(celsius) {
  farenheit <- (celsius * 9/5) + 32
  return(farenheit)
}

```

Use the function and check values
```{r}
farenheit <- celsius2fahr(celsius)
farenheit
farenheit == airtemps
```

