---
title: "Untitled"
author: "Lia Domke"
date: "11/4/2019"
output: html_document
---
This is a chunk that specifies how all the following chunks will be formated 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Different R markdown techniques
Some **bold** text. Have to include a blank line to start new paragraph

Include a [link](http://nceas.ucsb.edu)

Can create bulleted list:

* apple
    * pink lady
    * red delicious 
    * green
* banana
* orange

Can also include a numbered list, 
fruit in the order of priority: 

1. apple
    * pink lady 
    * red delicious
    * green
2. banana
3. orange

```{r, results = 'hide', echo = FALSE}
# Some information
x <- 17*2 
y <- 3*4

ans <- x/y
ans
```
## In-line R code
The results are `r round(ans, 3)`

## LaTeX  examples
$\sum{x^2}$

This equation "knowledge" is incorporated within R markdown however if you want to knit to pdf rather than html you will need to install a lite version of LaTex (package::tinytex). See help file from [link](https://learning.nceas.ucsb.edu/2019-11-RRCourse/introduction-to-r-and-rmarkdown.html#my-rmarkdown-wont-knit-to-pdf)

# Read in data
Use `read.csv` to read in tabular data

```{r}
# Using Rmarkdown the working directory will automatically be where the file folder where the script is stored
# This saves you from having to write the full directory. If you use relative paths (like this) it makes it easier 
# to share the working directory (i.e. folders etc) with someone else
# If you hit "tab" when you are in the parenthesis it will show you what is available within that wd
bg_chem <- read.csv("BGchem2008data.csv", stringsAsFactors = FALSE)
# stringsAsFactors argument makes it so that character vectors are NOT converted as factors
# can use `class` to see how different things are understood by R
class(bg_chem) 
# data frame - a list of named vectors (series of columns) that all have the same length 
# can look at a dataframe in a couples ways: can look at blue arrow in global environment, can view in 
# viewer - click on frame or View(name). When you look at this NOTHING in your object changes. 
```

## Look at the details of the data
```{r}
# use `head` to look at by default first 6 rows
head(bg_chem)
# use `mean` function of the CTD depth and save to variable
mean_depth <- mean(bg_chem$CTD_Depth)
# use base r plot to plot data
plot(bg_chem$CTD_Depth, bg_chem$CTD_Temperature)
```



